#Assignment 1A - Data Exploration
#IDS 572 | CRN 38886 | Fall 2021
#Team - Giovanni Alvin Prasetya, Karishma Mulchandani, and Rajaram Ramesh

#Libraries
library(tidyverse)
library(lubridate)
library(pROC)
library(broom)

#Reading Data File
lcdf <- read_csv('lcData100K.csv')

#Q2(a)(i)
```{r}
#1 - Loans shown by loan status
lcdf %>% group_by(loan_status) %>% summarise(nLoans=n())

#2 - Loans shown by loan status as a percentage of total loans
lcdf %>% group_by(loan_status) %>% summarise(nLoans=n()) %>% mutate(pct=nLoans/sum(nLoans) * 100)

#3 - Loans shown by loan status and grade (table)
table(lcdf$loan_status, lcdf$grade)

#4 - Loans shown by loan status and grade as a proportion (table)
LoanGrade <- table(lcdf$loan_status, lcdf$grade) 
ppLoanGrade <- prop.table(LoanGrade) 
ppLoanGrade <- ppLoanGrade*100
ppLoanGrade

#5 - Loans shown by loan status and grade as a proportion (barplot)
barplot(ppLoanGrade,
  main = "Loan Status by Grade", 
  xlab = "Proportion of Loans",
  ylab = "Grade",
  col = c("red", "green"),
  xlim = c(0,40), horiz = TRUE)
box (lwd=2)

#6 - Loans shown by loan status and sub-grade (table)
table(lcdf$loan_status, lcdf$sub_grade)

#7 - Loans shown by loan status and sub-grade as a proportion (table)

ppLoanSubGrade <- table(lcdf$loan_status, lcdf$sub_grade) 
ppLoanSubGrade <- prop.table(ppLoanSubGrade) 
ppLoanSubGrade <- ppLoanSubGrade*100 
ppLoanSubGrade

#8 - Loans shown by loan status and sub-grade as a proportion (barplot)
barplot(ppLoanSubGrade,
  main = "Loan Status by Sub-grade", 
  xlab = "Proportion of Loans",
  ylab = "Subgrade",
  col = c("red", "green"),
  xlim = c(0,8), horiz = TRUE)
box (lwd=2)

```

#Q2(a)(ii)
```{r}
#filter the Charged off and Fully paid
lcdf %>% group_by(loan_status, grade) %>% tally()
view(lcdata1)
lcdf1<-lcdata%>%select("loan_status","grade","sub_grade","loan_amnt")
lcdf2<- lcdata%>%select("loan_status","grade","sub_grade","loan_amnt","int_rate")

#using table
table(lcdata$loan_status, lcdata$grade)

#group by amount and loan
lcdf1 %>% group_by(grade) %>% summarise(sum(loan_amnt))
view(lcdf1)

#convert the lcdata1 list to dataframe
dataFrame <- as.data.frame(lcdf1)

#calculate loan amount by grades
a<-with(dataFrame, sum(loan_amnt[grade == 'A']))
a
b<-with(dataFrame, sum(loan_amnt[grade == 'B']))
b
c<-with(dataFrame, sum(loan_amnt[grade == 'C']))
c
d<-with(dataFrame, sum(loan_amnt[grade == 'D']))
d
e<-with(dataFrame, sum(loan_amnt[grade == 'E']))
e
f<-with(dataFrame, sum(loan_amnt[grade == 'F']))
f
g<-with(dataFrame, sum(loan_amnt[grade == 'G']))
g

#Variance of interest rate by grade and sub grade by average
lcdf2 %>% group_by(grade) %>% summarise(mean(int_rate))
lcdf2 %>% group_by(sub_grade) %>% summarise(mean(int_rate))

#Variance of average, std dev, min/max of interest rate by grade and subgrade
lcdf2 %>% group_by(grade) %>% summarise(averageInterest=mean(int_rate),stdevInterest=sd(int_rate), minInterest=min(int_rate), maxInterest=max(int_rate))
lcdf2 %>% group_by(sub_grade) %>% summarise(averageInterest=mean(int_rate),stdevInterest=sd(int_rate), minInterest=min(int_rate), maxInterest=max(int_rate))

#Plot the variance of loan amount by grade
ggplot(lcdf2, aes( x = loan_amnt)) + geom_histogram(aes(fill=grade))

#plot the variance of average interest rate base on grade and sub grade
ggplot(lcdf2, aes( x = int_rate)) + geom_histogram(aes(fill=grade))
ggplot(lcdf2, aes( x = int_rate)) + geom_histogram(aes(fill=sub_grade))

```

#Q2(a)(iii)
```{r}
getwd()
select('lcData100K.csv',c('loan_status'))
lcdf<-LoanData%>%select("loan_status","loan_amnt","grade","int_rate","last_pymnt_d","issue_d",
        "total_pymnt","funded_amnt","annRet") 
#%>%filter(loan_status=="Fully Paid")
view(lcdf)
lcdf$last_pymnt_d<-paste(lcdf$last_pymnt_d, "-01", sep = "")
#view(lcdf)
install.packages("lubridate")
library(lubridate)
lcdf$last_pymnt_d<-parse_date_time(lcdf$last_pymnt_d,"myd")
#head(lcdf[, c("last_pymnt_d", "issue_d")])

x<- as.duration(lcdf$issue_d  %--% lcdf$last_pymnt_d)/dyears(1)
view(lcdf)

#lcdf%>%filter(loan_status=="Fully Paid")

lcdf$actualTerm <- ifelse(lcdf$loan_status=="Fully Paid", 
                          as.duration(lcdf$issue_d  %--% lcdf$last_pymnt_d)/dyears(1), 3)
lcdf$actualReturn <- ifelse(lcdf$actualTerm>0,
                            ((lcdf$total_pymnt -lcdf$funded_amnt)/lcdf$funded_amnt)*(1/lcdf$actualTerm)*100, 0)

view(lcdf)

#lcdf %>% select(loan_status, int_rate, funded_amnt, total_pymnt, annRet, actualTerm, actualReturn) %>%  head()
#lcdf$testCol <- LoanData$annual_inc

boxplot(lcdf$actualTerm ~ lcdf$grade)

```

#Q2(a)(iv)
```{r}
#filter the Charged off and Fully paid
lcdf %>% group_by(loan_status, grade) %>% tally()
view(lcdata1)
lcdf1<-lcdata%>%select("loan_status","grade","sub_grade","loan_amnt")
lcdf2<- lcdata%>%select("loan_status","grade","sub_grade","loan_amnt","int_rate")

#using table
table(lcdata$loan_status, lcdata$grade)

#group by amount and loan
lcdf1 %>% group_by(grade) %>% summarise(sum(loan_amnt))
view(lcdf1)

#Calculate return from loan that labelled "charge off"
lcdf %>% group_by(loan_status) %>% summarise(sum(loan_status == "Charged Off"))

#expected return of loan amount from the int_rate
lcdf %>% select(loan_status, int_rate, funded_amnt, total_pymnt, grade) %>% head()

#Calculate annualized percentage return
lcdf$annRet <- ((lcdata$total_pymnt -lcdata$funded_amnt)/lcdata$funded_amnt)*(12/36)*100

#Summarise annual percentage return of "charged off" according to grade
lcdf %>% group_by(grade) %>% summarise(nLoans=n(), defaults=sum(loan_status=="Charged Off"), 
avgInterest= mean(int_rate), stdInterest=sd(int_rate), avgLoanAMt=mean(loan_amnt), avgPmnt=mean(total_pymnt), 
avgRet=mean(annRet), stdRet=sd(annRet), minRet=min(annRet), maxRet=max(annRet))

#Summarise annual percentage return of "charged off" according to sub grade
lcdf %>% group_by(sub_grade) %>% summarise(nLoans=n(), defaults=sum(loan_status=="Charged Off"), 
avgInterest= mean(int_rate), stdInterest=sd(int_rate), avgLoanAMt=mean(loan_amnt), avgPmnt=mean(total_pymnt), 
avgRet=mean(annRet), stdRet=sd(annRet), minRet=min(annRet), maxRet=max(annRet)) %>% view()

#Where do the negative numbers for minRet come from?
lcdf %>% select(loan_status, int_rate, funded_amnt, total_pymnt, annRet) %>% filter(annRet < 0) %>% head()

#are these all from 'Charged Off' loans?
lcdf %>% select(loan_status, int_rate, funded_amnt, total_pymnt, annRet) %>% filter(annRet < 0) %>% count(loan_status)

#Calculate the annual return percentage
lcdf$last_pymnt_d<-paste(lcdata$last_pymnt_d, "-01", sep = "")
lcdf$last_pymnt_d<-parse_date_time(lcdata$last_pymnt_d,  "myd")
lcdf$actualTerm <- ifelse(lcdata$loan_status=="Fully Paid", as.duration(lcdata$issue_d  %--% lcdata$last_pymnt_d)/dyears(1), 3)
lcdf$actualReturn <- ifelse(lcdata$actualTerm>0, ((lcdata$total_pymnt-lcdata$funded_amnt)/lcdata$funded_amnt)*(1/lcdata$actualTerm)*100, 0)
lcdf %>% select(loan_status, int_rate, funded_amnt, total_pymnt, annRet, actualTerm, actualReturn, sub_grade)%>% filter(loan_status == "Charged Off") %>% view()

```

#Q2(a)(v)
```{r}
lcdf$purpose <- LoanData$purpose
view(lcdf)

# Does default rate, int-rate, etc vary by loan purpose
lcdf %>% group_by(purpose) %>% tally()
lcdf %>% group_by(purpose) %>% summarise(nLoans=n(), defaults=sum(loan_status=="Charged Off"), defaultRate=defaults/nLoans, avgIntRate=mean(int_rate),  avgLoanAmt=mean(loan_amnt),  avgActRet = mean(actualReturn), avgActTerm=mean(actualTerm))

#Does loan-grade vary by purpose?
table(lcdf$purpose, lcdf$grade)

```

#Q2(a)(vi)
```{r}
#1 - Converting emp_length to factor, arranged in ascending number of years
lcdf$emp_length <- factor(lcdf$emp_length, levels = c("< 1 year","1 year","2 years","3 years","4 years","5 years","6 years","7 years","8 years","9 years","10+ years","n/a"))

#2 - Loans shown by employment length
lcdf %>% group_by(emp_length) %>% tally()

#3 - Loans shown by employment length and loan status
table(lcdf$loan_status, lcdf$emp_length)

#4 - Loans shown as a percentage of Charged Off loans for each level of employment length
cc = table(lcdf$loan_status, lcdf$emp_length)
(cc[1,]/(cc[1,] + cc[2,]))*100

#5 - Loans shown by employment length and loan grade
table(lcdf$grade, lcdf$emp_length)

#6 - Loans shown by employment length and loan purpose
table(lcdf$purpose, lcdf$emp_length)

#7 - Loans shown by employment length and home ownership status
table(lcdf$home_ownership, lcdf$emp_length)

```

#Q2(c)
```{r}
#1 - How many variables are there in the data file?
dim(lcdf)

#2 - Drop variables with all empty values
lcdf <- lcdf %>% select_if(function(x){!all(is.na(x))})

#3 - How many variables remain? 
dim(lcdf)

#4 - Initially we had 145 Variables, after running the code we kept 108 variables 

#5 - Missing value proportions showing only those columns where there are missing values 
colMeans(is.na(lcdf))[colMeans(is.na(lcdf))>0]

#6 - Missing value percentages showing only those columns where there are missing values 
colMeans(is.na(lcdf))[colMeans(is.na(lcdf))>0]*100

```

#Q3
```{r}
#Drop some columns that would not useful and those that would cause a leakage
lcdf <- lcdf %>% select(-c(funded_amnt_inv, term, emp_title, pymnt_plan, title, zip_code, addr_state, out_prncp, out_prncp_inv, total_pymnt_inv, total_rec_prncp, total_rec_int,total_rec_late_fee,recoveries, collection_recovery_fee, last_credit_pull_d, policy_code, disbursement_method, debt_settlement_flag, hardship_flag, hardship_dpd, settlement_term, application_type))

#To drop other variables,
#varsToRemove <- c("last_pymnt_d", "last_pymnt_amnt","annRet")
#lcdf <- lcdf %>% select(-varsToRemove)

```

#Q4
```{r}
#deploy aucAll variable considering both numeric and factor variable
aucAll<- sapply(lcdf %>% mutate_if(is.factor, as.numeric) %>% select_if(is.numeric), auc, response=lcdf$loan_status) 

#determine variable with auc > 0.5 and using tidy from broom package
aucAll[aucAll>0.5]
tidy(aucAll) %>% arrange(desc(aucAll))
tidy(aucAll[aucAll > 0.5]) %>% view()

```

